Data cleaning experiment 3

### 2023-11-08 Experiment_1/cleaning.Rmd
Remove cached *_files directories

```{r}
# recode convergence
d <- d %>%
  mutate(convergence = as.factor(convergence),
         convergence = fct_relevel(convergence, "noconv", "conv"),
         number = as.factor(number),
         number = fct_relevel(number, "small", "large"))
```

>>>>

```{r}
# recode convergence
d <- d %>%
  mutate(
    convergence = ifelse(convergence == "conv", "convergent", "divergent"),
    convergence = as.factor(convergence),
    convergence = fct_relevel(convergence, "divergent", "convergent"),
    number = as.factor(number),
    number = fct_relevel(number, "small", "large"))
```

### 2023-11-08 Experiment_1/model.Rmd
Remove cached *_files directories

```{r, message=FALSE, echo=FALSE}
# Load packages
library(tidyverse)
library(gghalves) # for plots
library(patchwork)
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
```

```{r, echo = FALSE}
#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

@@

```{r}
# Imagine a situation in which answers range from 1000 to 2000
range <- 1000
# and a population of n = 999
population <- 999
# and we observe samples of 3
sample <- 3
```

@@

We suppose that these competence levels are drawn from a random uniform distribution. That is, in a population, each level of competence is equally likely to occur.

@@

```{r}
# 1: Distribution of competence
data <- data.frame(competence = runif(population, 
                                      min = range/1000, 
                                      max = range)) %>% 
  # Scale competence to reach from 0 to 1 (instead of 1 to 1000).
  # Reverse the scale such that 1 becomes maximal competence (instead of 0).
  mutate(competence_scaled = 1 - ((competence - 1) / (1000-1)))

# generated population
ggplot(data, aes(x = competence_scaled)) +
  geom_histogram() + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```
### Generate a choice for each individual based on their competence.

The numbers that we generate from the normal distributions are tuncated such that they all lie within the pre-defined range (1000 to 2000)

@@

data$answer <- data$competence %>% 
  purrr::map_dbl(function(x){

@@

    answer = truncnorm::rtruncnorm(1, mean = mean, sd = sd, a = 1000, b = 2000)
  }
  )

@@

We measure accuracy as the (absolute value of) distance between the chosen answer and the true answer for each individual.

@@

```{r}
# 3: measure accuracy
data <- data %>% 
  mutate(accuracy = -1 * abs(mean - answer))

@@

ggplot(data, aes(x = competence_scaled, y = accuracy)) +
  geom_point()
```

@@

We define convergence as the standard deviation of the answers within a sample (multiplied by -1 so that higher values correspond to greater convegence)

@@

# compute the summary statistics for each sample
results <- data %>% 
  group_by(sample_id, convergence) %>% 
  summarize(competence_mean = mean(competence_scaled),
            accuracy_mean = mean(accuracy)) %>% 
  # store simulation information
  mutate(population = population, 
         sample = sample)
```

```{r}
accuracy_plot <- ggplot(results, aes(x = convergence, y = accuracy_mean)) +
  geom_point() +
  # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Accuracy \n(average distance to true mean)") +
  plot_theme

competence_plot <- ggplot(results, aes(x = convergence, y = competence_mean)) +
  geom_point() +
  # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Competence \n(SD for data generating function, transposed)") +
  plot_theme

@@

```{r, echo=FALSE}
simulate_single_population <- function(population, sample, min_range, max_range, 
                                       competence_ratio = 1000) {

@@

  # 1: Distribution of competence
  data <- data.frame(competence = runif(population, 
                                        min = range/competence_ratio, 
                                        max = range)) %>% 
    # Scale competence to reach from 0 to 1.
    # Reverse the scale such that 1 becomes maximal competence (instead of 0).
    mutate(competence_scaled = 1 - ((competence - range/competence_ratio) / 
                                      (range - range/competence_ratio)))

@@

  data$answer <- data$competence %>% 
    purrr::map_dbl(function(x){

@@

  # 3: measure accuracy
  data <- data %>% 
    mutate(accuracy = -1 * abs(mean - answer))

@@

  # 6: calculate accuracy and competence levels per sample
  results <- data %>% 
    group_by(sample_id, convergence) %>% 
    summarize(competence_mean = mean(competence_scaled),
              accuracy_mean = mean(accuracy)) %>% 
    # store simulation information
    mutate(population = population, 
           sample = sample) %>% 
    ungroup()

  return(results)
}

```{r, echo=FALSE}
# # # test output (set return to data above)
# test <- simulate_single_population(population = 500, sample = 3,
#                                    min_range = 1, max_range = 6,
#                                    competence_ratio = 1000)
```

@@

```{r, echo=FALSE}
# function for different sample sizes
vary_sample <- function(file_name, n, ...) {

@@

@@

  # extract simulation info
  simulation_info <- d %>% 
    summarize(across(c(population, sample, total_iterations), max)) %>% 
    pivot_longer(everything(), names_to = "parameter", values_to = "value") %>% 
    gridExtra::tableGrob(cols = NULL, rows = NULL)

  plot_accuracy <- ggplot(d, aes(x = convergence, y = accuracy_mean)) +
    geom_hex() +
    # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Accuracy \n(average distance to true mean)") +
    plot_theme +
  scale_fill_viridis_c(option = "plasma", begin = 0.1)

  plot_competence <- ggplot(d, aes(x = convergence, y = competence_mean)) +
    geom_hex() +
    # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Competence \n(SD for data generating function, transposed)") +
    plot_theme +
  scale_fill_viridis_c(option = "plasma", begin = 0.1)

@@

    return((plot_accuracy | plot_competence) / simulation_info)
  }

@@

```{r, echo=FALSE}
plot_results_vary <- function(data, outcome = "everything", sample) {

@@

    sample_plot <- plot_results(data = plot_data)

@@

@@

### Generate data

@@

# run simulation and store results in .csv files
vary_sample(population = 999,
            min_range = 1000, max_range = 2000,
            competence_ratio = 1000,
            iterations = 100,
            file_name = "model_vary_3_10_20_50.csv",
            n = n)

# read simulated data from .csv files
data <- read_csv("data/model_vary_3_10_20_50.csv")
```

@@

```{r}
# plot results
plot_results_vary(data)
```

@@

Below, what I did was predict accuracy and competence for convergence levels of 20 and 150 based on a regression run on the model data. We could compare that to the participant data - but it's not quite the same thing...

@@

>>>>

```{r, message=FALSE, echo=FALSE}
# Load packages
library(tidyverse)
library(gghalves) # for plots
library(patchwork)
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
library(ggpubr) # for plots and ggarrange()
```

```{r, echo = FALSE}
# set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold title
    plot.title = element_text(face = "bold", size = rel(0.7)),
    # Plain subtitle that is grey
    plot.subtitle = element_text(face = "plain", size = rel(1), color = "grey70"),
    # Bold legend titles
    legend.title = element_text(face = "bold"),
    # Bold axis titles
    axis.title = element_text(face = "bold"), 
    # Bold legend labels
    legend.text = element_text(face = "bold")
    )
```

@@

```{r}
# Imagine a situation in which answers range from 1000 to 2000
range <- 1000
# and a population of n = 999
population <- 999
# and we observe samples of 3
sample <- 3
# alpha and beta values for beta distributions modelling competence
alpha <- 1
beta <- 1
# ramge of values for estimates
min_range = 1000
max_range = 2000

# set arbitrary limit for competence distribution
# (factor determining distance between lowest and highest possible competence, 
# i.e. standard deviation)
competence_ratio = 1000 
```

@@

We assume that individuals (i) vary in competence and (ii) that the distribution of competence in the population varies, as can be modeled by a beta distribution. In the case below, we assume a uniform competence distribution (equal to a beta distribution of alpha = 1 and beta = 1).

@@

```{r}
# 1: Distribution of competence

# Theoretical population distribution of competence
ggplot() +
  geom_function(fun = ~dbeta(.x, shape1 = 1, shape2 = 1))

# randomly draw competence levels
data <- tibble(id = 1:population,
               competence = rbeta(population, shape1 = alpha, 
                                          shape2 = beta)) %>% 
    # For now competence tending towards one is more competent.
    # For the data generating, we need the reverse and scale this competence measure.  
    mutate(
      # Since competence = standard deviation of estimate generating normal distribution, 
      # reverse the scale such that 0 corresponds to maximal competence (instead of 1).
      competence_reversed = 1 - competence, 
      # Put values on a scale: 
      # We (arbitrarily) set the lowest competence to the range of possible values, 
      # in case of our experiments 2000 - 1000 = 1000. 
      # We set the highest competence to 1/1000 of the range of possible values, 
      # in our case 1.
      competence_reversed_scaled = (competence_reversed * range) + range/competence_ratio)

# generated population
ggplot(data, aes(x = competence)) +
  geom_histogram() + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```
### Generate a choice for each individual based on their competence.

The numbers that we generate from the normal distributions are truncated such that they all lie within the pre-defined range (1000 to 2000)

@@

# use the reversed competence variable to generate the estimates
data$answer <- data$competence_reversed_scaled %>% 
  purrr::map_dbl(function(x){

@@

@@

We measure accuracy as the (squared) distance between the chosen answer and the true answer for each individual.

@@

```{r}
# 3: measure accuracy
data <- data %>% 
  mutate(accuracy = -1 * (mean - answer)^2)

@@

ggplot(data, aes(x = competence, y = accuracy)) +
  geom_point()
```

@@

We define convergence as the standard deviation of the answers within a sample (multiplied by -1 so that higher values correspond to greater convergence)

@@

# compute the summary statistics for each sample
results <- data %>% 
  group_by(sample_id, convergence) %>% 
  summarize(competence_mean = mean(competence),
            accuracy_mean = mean(accuracy)) %>% 
  # store simulation information
  mutate(population = population, 
         sample = sample)
```

```{r}
accuracy_plot <- ggplot(results, aes(x = convergence, y = accuracy_mean)) +
  geom_point() +
  # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Accuracy \n(squared distance to true mean)") +
  plot_theme

competence_plot <- ggplot(results, aes(x = convergence, y = competence_mean)) +
  geom_point() +
  # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Competence \n(SD of data generating function)") +
  plot_theme

@@

```{r, echo=FALSE}
simulate_single_population <- function(population, sample, min_range, max_range, 
                                       competence_ratio = 1000, alpha, beta) {

@@

  # 1: Distribution of competence
  data <- tibble(id = 1:population,
                 competence = rbeta(population, shape1 = alpha, 
                                    shape2 = beta)) %>% 
    # For now competence tending towards one is more competent.
    # For the data generating, we need the reverse and scale this competence measure.  
    mutate(
      # Since competence = standard deviation of estimate generating normal distribution, 
      # reverse the scale such that 0 corresponds to maximal competence (instead of 1).
      competence_reversed = 1 - competence, 
      # Put values on a scale: 
      # We (arbitrarily) set the lowest competence to the range of possible values, 
      # in case of our experiments 2000 - 1000 = 1000. 
      # We set the highest competence to 1/1000 of the range of possible values, 
      # in our case 1.
      competence_reversed_scaled = (competence_reversed * range) + range/competence_ratio)

@@

  data$answer <- data$competence_reversed_scaled %>% 
    purrr::map_dbl(function(x){

@@

  # 3: measure accuracy
  data <- data %>% 
    mutate(accuracy = -1 * (mean - answer)^2)

@@

  # 6: calculate accuracy and competence levels per sample
  results <- data %>% 
    group_by(sample_id, convergence) %>% 
    summarize(competence_mean = mean(competence),
              accuracy_mean = mean(accuracy)) %>% 
    # store simulation information
    mutate(population = population, 
           sample = sample) %>% 
    ungroup()

  return(results)
}
```

```{r, echo=FALSE}
# test <- simulate_single_population(population = 500, sample = 3,
#                                    min_range = 1, max_range = 6,
#                                    competence_ratio = 1000,
#                                    alpha = 1,
#                                    beta = 1)
```

@@

```{r, echo=FALSE}
# function for different sample sizes
vary_sample <- function(file_name, n, format = "result", ...) {

@@

    if(format == "result") {
      return(data)
    } else{
      write_csv(data, paste0("data/", file_name))
    }
  }
}
```

#### Vary competence

We now want to do the above (simulate a single population, repeat this process, do so for various sample sizes and choice options) for different competence distributions. We vary these distributions with the `alpha` and `beta` parameters for the beta distributions.

```{r, echo=FALSE}
# set name
file_name <- "model.csv" # change for new analyses / or delete file to re-use same name

vary_competence <- function(file_name, alpha, beta, names, ...) {

    competence_data <- tibble(alpha, beta, names) %>% 
      mutate(competence = paste0(names, "\n(alpha = ", alpha, ", beta = ", beta, ")"))

    # do all the above functions for different competence distributions
    data <- competence_data$competence %>% purrr::map_df(function(competence_level){
      # this is essentially a for loop

      # To keep track of progress
      print(paste("competence_level = ", competence_level))

      # set alpha and beta
      alpha = competence_data %>% filter(competence == competence_level) %>% pull(alpha)
      beta = competence_data %>% filter(competence == competence_level) %>% pull(beta)

      # run simulation
      result <- vary_sample(
        alpha = alpha, beta = beta, file_name = file_name, ...)

      # add competence level
      result <- result %>% mutate(competence = competence_level, 
                                  alpha = alpha,
                                  beta = beta)

@@

  # extract simulation info
  caption = paste0("Sample size of informants: ", d$sample[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])

  plot_accuracy <- ggplot(d, aes(x = convergence, y = accuracy_mean)) +
    geom_hex() +
    # Add nice labels
    labs(x = "Convergence \n(SDs of samples)", 
         y = "Accuracy \n(squared distance to true mean)", 
         caption = paste0(caption, "; assuming uniform competence distribution")) +
    plot_theme +
  scale_fill_viridis_c(option = "plasma", begin = 0.1)

  plot_competence <- ggplot(d, aes(x = convergence, y = competence_mean)) +
    geom_hex() +
    # Add nice labels
    labs(x = "Convergence \n(SDs of samples)", 
         y = "Competence \n(SD for data generating function)",
         caption = paste0(caption, "; assuming uniform competence distribution")) +
    plot_theme +
  scale_fill_viridis_c(option = "plasma", begin = 0.1)

@@

    return(plot_accuracy | plot_competence)
  }

@@

```{r, echo=FALSE}
plot_results_vary <- function(data, ...) {

@@

    sample_plot <- plot_results(data = plot_data, ...)

@@

#### c) Plot competence distributions

We want to plot the underlying competence distributions.

```{r}
# Plot competence distributions
plot_competence_distributions <- function(data) {
  # Your data frame
  d <- data %>% 
    group_by(competence) %>% 
    reframe(alpha = unique(alpha), beta = unique(beta))

  # Function to plot the different competence distributions 
  plot_competence_distributions <- function(alpha, beta) {
    ggplot() +
      geom_function(
        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
      ) +
      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
           y = "Density", x = "Competence") +
      plot_theme
  }

  # Specify the values of alpha and beta from the data
  alpha_values <- d$alpha
  beta_values <- d$beta

  # Create a list of ggplot objects
  plot_list <- map2(alpha_values, beta_values, plot_competence_distributions)

  # Arrange and display the plots
  competence_plot <- ggarrange(plotlist = plot_list) %>%
    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
                                                    face = "bold",
                                                    color = "darkgrey",
                                                    size = 14))

  return(competence_plot) 
}
```

#### d) Plot varying competence distributions by sample

Fourth, we want another function that plots results by various competence levels by sample sizes.

```{r}
plot_competence <- function(data, outcome = "accuracy") {

  # extract simulation info
  caption = paste0("Sample size of informants: ", d$sample[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])

  plot_accuracy <- ggplot(d, aes(x = convergence, y = accuracy_mean)) +
    geom_hex(alpha = 0.9) +
    #geom_smooth(method = "lm", se = FALSE) +
    stat_smooth(geom='line', method = "lm", alpha=0.7, se=FALSE, color = "darkgrey") +
    # Add nice labels
    labs(x = "Convergence \n(SDs of samples)", 
         y = "Accuracy \n(squared distance to true mean)",
         caption = caption) +
  scale_fill_viridis_c(option = "plasma", begin = 0.1) +
    facet_wrap(~competence) +
    plot_theme + 
    theme(legend.position = "top", 
          legend.key.width = unit(1.5, "cm"), 
          legend.key.height = unit(0.3, "cm"))

  plot_competence <- ggplot(d, aes(x = convergence, y = competence_mean)) +
    geom_hex(alpha = 0.9) +
    #geom_smooth(method = "lm", se = FALSE) +
    stat_smooth(geom='line', method = "lm", alpha=0.7, se=FALSE, color = "darkgrey") +
    # Add nice labels
    labs(x = "Convergence \n(SDs of samples)", 
         y = "Competence \n(SD for data generating function)",
         caption = caption) +
  scale_fill_viridis_c(option = "plasma", begin = 0.1) +
    facet_wrap(~competence) +
    plot_theme + 
    theme(legend.position = "top", 
          legend.key.width = unit(1.5, "cm"), 
          legend.key.height = unit(0.3, "cm"))

    return(plot_competence)
  }
}

plot_competence_vary <- function(data, ...) {

    sample_plot <- plot_competence(data = plot_data, ...)

  return(plot_list)
}

plot_competence_vary_relative_majority <- function(data, variable = options, outcome = "Accuracy") {

  # Extract variable name as a string
  variable_name <- deparse(substitute(variable))

  # retrieve info from choice of variable
  if (variable_name == "options") {
    caption = paste0("Sample size of informants: ", d$sample[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])
  }
  if (variable_name == "sample") {
    caption = paste0("Number of choice options: ", d$options[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])
  }

  d <- d %>% 
    group_by(relative_majority, {{variable}}, competence) %>%
    summarise(competence_mean = mean(average_relative_competence),
              accuracy_mean = mean(average_accuracy),
              accuracy_sd = sd(average_accuracy),
              competence_sd = sd(average_relative_competence)
    )

  # Assign outcome variable
  if(outcome == "Accuracy") {
    d$outcome <- d$accuracy_mean
    d$outcome_sd <- d$accuracy_sd
  }

  if(outcome == "Competence") {
    d$outcome <- d$competence_mean
    d$outcome_sd <- d$competence_sd
  }

  plot_outcome <- ggplot(d, 
                         aes(x = relative_majority, y = outcome,
                             color = as.factor({{variable}}))) +
    geom_point() +
    geom_line() +
    # Add nice labels
    labs(x = "Relative majority \n (share of informants agreeing on an option)", 
         y = outcome, caption = caption, 
         color = variable_name) +
    scale_color_viridis_d(option = "plasma", begin = 0.1, 
                          #limits = rev(levels(d$constellation)),
                          direction = -1
    ) +
    facet_wrap(~competence) +
    plot_theme + 
    theme(legend.position = "top")

   return(plot_outcome)

@@

### Experiment 4

#### Generate Data

@@

# run simulation and store results in .csv files
vary_sample(population = 999,
            min_range = 1000, max_range = 2000,
            competence_ratio = 1000,
            iterations = 100,
            file_name = "sim_experiment_1.csv",
            n = n, 
            alpha = 1, 
            beta = 1, 
            format = ".csv")

# read simulated data from .csv files
data <- read_csv("data/sim_experiment_1.csv")
```

@@

```{r}
# plot results
plot_results_vary(data, outcome = "accuracy")
plot_results_vary(data, outcome = "competence")
```

@@

Below, what I did was predict accuracy and competence for convergence levels of 20 and 150 based on a regression run on the model data. We could compare that to the participant data - but it's not quite the same thing..

@@

## Vary competence

```{r, message=FALSE}
n <- c(3, 5, 10, 20, 50, 100)
alpha = c(0.1, 2, 1, 1, 6, 0.9)
beta = c(1, 6, 1, 0.1, 2, 0.9)
names = c("very incompetent", "incompetent", "uniform", 
          "very competent", "competent", "bimodal")

# run simulation and store results in .csv files
vary_competence(population = 999,
            min_range = 1000, max_range = 2000,
            competence_ratio = 1000,
            iterations = 100,
            file_name = "sim.csv",
            n = n, 
            alpha = alpha, 
            beta = beta, 
            names = names)

# read simulated data from .csv files
test <- read_csv("data/sim.csv")
```

```{r}
plot_competence_distributions(test)
```
```{r}
# for accuracy
plot_competence_vary(data = test, outcome = "accuracy")
```
```{r}
# for competence
plot_competence_vary(data = test, outcome = "competence")
```

### 2023-11-08 Experiment_3/cleaning.Rmd
Remove cached *_files directories

Lastly, we recode the `convergence` variable. For now, the baseline category is _convergence_, which will yield a negative estimate for the effect of _no convergence_. We want to have the positive effect of _convergence_ instead. Likewise, we recode the levels of the `independence` variable to make sure the effect goes into the right direction (i.e. the effec of _independence_ compared to _conflict of interest_).

```{r}
d <- d %>%
  mutate(convergence = as.factor(convergence),
         convergence = fct_relevel(convergence, "div", "conv"),
         independence = as.factor(independence),
         independence = fct_relevel(independence, "confl", "indep")
         )
```

>>>>

Lastly, we recode the `convergence` variable. For now, the baseline category is _convergence_, which will yield a negative estimate for the effect of _no convergence_. We want to have the positive effect of _convergence_ instead. Likewise, we recode the levels of the `independence` variable to make sure the effect goes into the right direction (i.e. the effect of _independence_ compared to _conflict of interest_).

```{r}
d <- d %>%
  mutate(
    convergence = ifelse(convergence == "conv", "convergent", "divergent"),
    convergence = as.factor(convergence),
    convergence = fct_relevel(convergence, "divergent", "convergent"),
    independence = ifelse(independence == "indep", "independent", "conflict"),
    independence = as.factor(independence),
    independence = fct_relevel(independence, "conflict", "independent")
         )
```
